Peer Assessment_1 W2_M5
========================================================
Read de data, in the working directory, and modify the time format

```{r}
data <- read.csv("activity.csv", header=T)
data$date <- as.Date(data$date, "%Y-%m-%d")
```
## QUESTION No.1
### What is mean total number of steps taken per day?
**For this part of the assignment, you can ignore the missing values in the dataset**

**1. Make a histogram of the total number of steps taken each day**

```{r}
sum.day.deletingNA    <- with(subset(data, steps != "NA"), 
                              tapply(steps, date, sum))
mean.day.deletingNA   <- with(subset(data, steps != "NA"), 
                              tapply(steps, date, mean))
median.day.deletingNA <- with(subset(data, steps != "NA"), 
                              tapply(steps, date, median))
```
```{r fig.width=7, fig.height=6}
hist(sum.day.deletingNA, col="green", 
     main="Histogram of the Total Number of Steps Taken Each Day",
     xlab="Number of steps taken each day")
mtext("--- deleting NA data ---")
```

**2. Calculate and report the mean and median total number of steps taken**
**per day**

```{r}
 (rpta.deletingNA <- data.frame(DATE=names(mean.day.deletingNA), 
                               MEANS.DAY.delNA=as.vector(mean.day.deletingNA), 
                               MEDIANS.DAY.delNA=as.vector(median.day.deletingNA)))
```

## QUESTION No.2
### What is the average daily activity pattern?

**1. Make a time series plot (i.e. type = "l" ) of the 5-minute interval (x-axis)**
**and the average number of steps taken, averaged across all days (y-axis)**

```{r}
mean.interval <- with(subset(data, steps != "NA"), tapply(steps, interval, mean))
```
```{r fig.width=7, fig.height=6}
plot(names(mean.interval), as.vector(mean.interval), type="l", col=4,    
     xlab="interval", ylab="Average number of steps", main="Time Series")
```

**2. Which 5-minute interval, on average across all the days in the dataset,**
**contains the maximum number of steps?**

```{r}
(names(which(mean.interval==max(mean.interval))))
```

## QUESTION No.3
### What is the impact of imputing missing data on the estimates of the total
### daily number of steps?

**1. Calculate and report the total number of missing values in the dataset**
 (i.e. the total number of rows with NAs)
```{r}
sum(is.na(data$steps))
```

**2. Devise a strategy for filling in all of the missing values in the dataset.**
The strategy does not need to be sophisticated. For example, you could use
the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r}
steps.imputingNA <- ifelse(is.na(data$steps), 
                           mean.interval[as.character(data$interval[which(is.na(data$steps))])],
                           data$steps)
```

**3. Create a new dataset that is equal to the original dataset but with the**
**missing data filled in**
```{r}
data.imputingNA       <- data
data.imputingNA$steps <- steps.imputingNA
```

**4. Make a histogram of the total number of steps taken each day and Calculate**
**and report the mean and median total number of steps taken per day.**
```{r}
sum.day.imputingNA    <- with(data.imputingNA, tapply(steps, date, sum))
mean.day.imputingNA   <- with(data.imputingNA, tapply(steps, date, mean))
median.day.imputingNA <- with(data.imputingNA, tapply(steps, date, median))
```
```{r fig.width=7, fig.height=6}
hist(sum.day.imputingNA,   col="blue",
     main="Histogram of the Total Number of Steps Taken Each Day",
     xlab="Number of steps taken each day")
mtext("--- imputing NA data ---")
```
```{r}
(rpta.imputingNA <- data.frame(DATE=names(mean.day.imputingNA), 
                               MEANS.DAY.impNA=as.vector(mean.day.imputingNA), 
                               MEDIANS.DAY.impNA=as.vector(median.day.imputingNA)))
```

## Do these values differ from the estimates from the first part of the assignment?
**----------------- Visualization**
```{r fig.width=7, fig.height=6}
par(mfrow=c(2,3))
par(mar=c(5.1, 4.1, .1, 2.1))
par(oma=c(0, 0, 4, 0))

hist(sum.day.deletingNA, col="green", xlab="", main="", 
     ylab="deleting NA data")
hist(mean.day.deletingNA, col="green", main="", ylab="", xlab="")
hist(median.day.deletingNA, main="", ylab="", xlab="")
hist(sum.day.imputingNA, col="blue", main="", xlab="Total per day",
     ylab="imputing NA data")
hist(mean.day.imputingNA, col="blue", main="", ylab="", xlab="Mean per day")
hist(median.day.imputingNA, col="blue", main="", ylab="", xlab="Median per day")
title(main="Number of Steps Taken Each Day", outer=T)
```

**----------------- Tabular comparison**
```{r}
summaryComparison <- function(vector1, vector2){
        a      <- summary(vector1)
        namesT <- names(a)
        a      <- as.vector(a)
        b      <- as.vector(summary(vector2))
        summaryMeans <- data.frame(a, b)
        rownames(summaryMeans) <- namesT
        colnames(summaryMeans) <- c("Vector 1", "Vector 2")
        summaryMeans
}

summaryComparison(mean.day.deletingNA, mean.day.imputingNA)
```


## QUESTION No.4
### Are there differences in activity patterns between weekdays and week-ends?
**For this part the weekdays() function may be of some help here. Use the dataset**
**with the filled-in missing values for this part**

**1. Create a new factor variable in the dataset with two levels - "weekday"**
**and "weekend" indicating whether a given date is a weekday or weekend day**
```{r}
we           <- c("Monday"=1, "Tuesday"=2, "Wednesday"=3, "Thursday"=4, 
                  "Friday"=5, "Saturday"=6, "Sunday"=7)
wee          <- weekdays(data.imputingNA$date)
week         <-we[wee]
data.imputingNA$week <- as.factor(ifelse(week<6,"weekday","weekend"))
```

**Make a panel plot containing a time series plot (i.e. type = "l" ) of the**
**5-minute interval (x-axis) and the average number of steps taken, averaged**
**across all weekday days or weekend days (y-axis).**
```{r}
meanIntervalWeekday <- with(subset(data.imputingNA, week=="weekday"), 
                            tapply(steps, interval, mean))

meanIntervalWeekend <- with(subset(data.imputingNA, week=="weekend"), 
                            tapply(steps, interval, mean))
```


**----------------- Visualization**
```{r fig.width=7, fig.height=6}
par(mfrow=c(3,1))
par(mar=c(5.1, 4.1, .1, 2.1))
par(oma=c(0, 0, 4, 0))
plot(names(meanIntervalWeekday), as.vector(meanIntervalWeekday), type="l", col=4,    
     xlab="", ylab="Weekday")
plot(names(meanIntervalWeekend), as.vector(meanIntervalWeekend), type="l", col=2,    
     xlab="", ylab="Weekend")
plot(names(meanIntervalWeekday), as.vector(meanIntervalWeekday), type="l", col=4,    
     xlab="interval", ylab="", main="")
lines(names(meanIntervalWeekend), as.vector(meanIntervalWeekend), type="l", col=2,    
      xlab="interval", ylab="Average number of steps")
legend("topright", lty=1, col=c(4, 2),
       legend=c("Weekday", "Weekend"))
title(main="Average Number of Steps Taken Each Interval", outer=T)
```

**----------------- Tabular comparison**
```{r}
summaryComparison(meanIntervalWeekday, meanIntervalWeekend)
```
